HealpixNet: Efficient spherical Convolutional Neural Network with HEALPix sampling for cosmological applications

### 2018-08-10 scnn.tex
todo: font
avoid clash of notation: loss function => cost function

\section*{TODO list}
\begin{itemize}[noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt]
	\item (TK) Related work in cosmo. The only cosmological applications of spherical CNNs in the related work are about projection on the tangent plane \citep{schmelze2017cosmologicalmodel, fluri2018deep, gupta2018nongaussianinformation}. Aren't there applications in cosmology for the other (c.f. related work section) approaches? Most of them are illustrated in omnidirectional imagery.
	\item (MD) distinction between CNN and FCN
    \item (first version done!)(NP+MD) paragraph on our chosen architecture in the Experiments sections
    \item (Done) (NP) vertex interpretation
    \item (MD) Are you happy with the vertex interpretation in section 2.5 Efficient convolutions?
    \item (NP) Finalise readme on Github code
    \item branch with notebooks with plots (results) NP
    \item NP move the repo to HealpixNet
    \item check if papers which appear as arxiv were not published in a conference or journal
    \item TK check part sphere notebook and whole sphere
    \item (done!) (TK) update the README for data zip - file from NP
    \item (almost done) TK read everything
    \item (done!) (TK) add small description of maps\_downsampled\_64.npz in the whole\_sphere notebook. \nati{We need to talk about that.}
    \item (done!) (TK) re-make figures to make the labels/legend/titles the same font size as the main text
    \item (done!) (TK) add info on comparison with histograms, so that it's just a demonstration
    \item (done!) (TK) add reference to figure (pooling) to the area of orders
    \item (done!) (TK) add section in conclusions/ paragraph on possible extensions to the code: dense output, regression, GANs, segmentation
	\item Beware to not confuse reader with synonyms, e.g. pixelization and sampling.
    \item Check names: HealpixNet, HealpixCNN, HealpixFCN
    \item Be sure that the convolution on SO(3) is clear: we want to say (a) is the one associated with the SO(3) rotation group (b) can be performed  in the spectral domain by a multiplication (c) is equivariant to rotation (d) is described well in Section 2 of \citep{kondor2018clebsch}.
	\item Have a "Contributions" paragraph?
	\item Notation: write element $i$ of vector $y$ as $y_i$, $y(i)$, or $y[i]$? Be consistent. Same for matrices.
	\item Try to make the paper shorter?
\end{itemize}

@@

The loss function $\ell(\Y, \bar \Y) = \ell(NN_\theta(\X), \bar \Y)$ measures how good the prediction $\Y$ is for sample $\X$, given the ground truth $\bar \Y$. For a classification task, the loss is usually taken to be the cross-entropy
\begin{equation*}
	\ell(\Y, \bar \Y) = - \sum_{i=1}^{N_{pix}} \sum_{j=1}^{N_{classes}} \bar y_{ij} \log(y_{ij}),
\end{equation*}
where $\bar \Y \in \R^{N_{pix} \times N_{classes}}$ is the ground truth label indicator, i.e., $\bar y_{ij} = 1$ if pixel $i$ of sample $\X$ belongs to class $j$ and is zero otherwise. For global prediction, the loss is as if $N_{pix} = 1$.
For a regression task, a common choice is the mean squared error (MSE)
\begin{equation*}
	\ell(\Y, \bar \Y) = \frac{\|\Y-\bar{\Y}\|_2^2}{N_{pix} F_{out}} = \frac{1}{N_{pix} F_{out}} \sum_{i=1}^{N_{pix}} \sum_{j=1}^{F_{out}} (y_{ij} - \bar y_{ij})^2,
\end{equation*}
where $\bar{\Y}$ is the desired output. Again, take $N_{pix} = 1$ for global regression. We emphasize that the loss and the SM layer is the sole difference between a neural network engineered for classification or regression.

The goal of learning is to find the parameters $\theta$ of the neural network that minimize the risk $R(\theta) = \esp \left[ \ell \left( NN_\theta(\X), \bar \Y \right) \right]$. In general, that expectation cannot be computed as the data distribution is unknown. We can however minimize an approximation, the empirical risk over the training set $\left\{ \left( \X_i, \bar \Y_i \right) \right\}_{i=1}^{N_{samples}}$:
\begin{equation*}
	\hat{\theta} = \argmin_\theta \sum_{i=1}^{N_{samples}} \ell \left(NN_\theta(\X_i), \bar \Y_i \right).
\end{equation*}
The optimization is performed by computing an error gradient w.r.t.\ all the parameters by back-propagation and updating them with a form of stochastic gradient descent (SGD):
\begin{equation*}
	\theta \leftarrow \theta - \frac{\eta}{|\B|} \sum_{i \in \B} \frac{\partial \ell \left( NN_\theta(\X_i), \bar \Y_i \right)}{\partial \theta} ,
\end{equation*}
where $\eta$ is the learning rate, and $\B$ is the set of indices in a mini-batch. Batches are used instead of single samples to gain speed by exploiting the parallelism afforded by modern computing platforms.

>>>>

\section*{TODO list}
\begin{itemize}[noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt]
	\item (TK) Related work in cosmo. The only cosmological applications of spherical CNNs in the related work are about projection on the tangent plane \citep{schmelze2017cosmologicalmodel, fluri2018deep, gupta2018nongaussianinformation}. Aren't there applications in cosmology for the other (c.f. related work section) approaches? Most of them are illustrated in omnidirectional imagery.
	\item (MD) distinction between CNN and FCN
    \item (first version done!)(NP+MD) paragraph on our chosen architecture in the Experiments sections
    \item (Done) (NP) vertex interpretation
    \item (MD) Are you happy with the vertex interpretation in section 2.5 Efficient convolutions?
    \item (NP) Finalise readme on Github code
    \item branch with notebooks with plots (results) NP
    \item NP move the repo to HealpixNet
    \item check if papers which appear as arxiv were not published in a conference or journal
    \item TK check part sphere notebook and whole sphere
    \item (done!) (TK) update the README for data zip - file from NP
    \item (almost done) TK read everything
    \item (done!) (TK) add small description of maps\_downsampled\_64.npz in the whole\_sphere notebook. \nati{We need to talk about that.}
    \item (done!) (TK) re-make figures to make the labels/legend/titles the same font size as the main text
    \item (done!) (TK) add info on comparison with histograms, so that it's just a demonstration
    \item (done!) (TK) add reference to figure (pooling) to the area of orders
    \item (done!) (TK) add section in conclusions/ paragraph on possible extensions to the code: dense output, regression, GANs, segmentation
	\item Beware to not confuse reader with synonyms, e.g. pixelization and sampling.
    \item Check names: HealpixNet, HealpixCNN, HealpixFCN
    \item Be sure that the convolution on SO(3) is clear: we want to say (a) is the one associated with the SO(3) rotation group (b) can be performed  in the spectral domain by a multiplication (c) is equivariant to rotation (d) is described well in Section 2 of \citep{kondor2018clebsch}.
	\item Have a "Contributions" paragraph?
	\item Notation: write element $i$ of vector $y$ as $y_i$, $y(i)$, or $y[i]$? Be consistent. Same for matrices.
	\item Try to make the paper shorter?
	\item Font: do we have to use times (set in the documentclass)? I (MD) find the rendering of math a bit ugly.
\end{itemize}

@@

The cost (or loss) function $C(\Y, \bar \Y) = C(NN_\theta(\X), \bar \Y)$ measures how good the prediction $\Y$ is for sample $\X$, given the ground truth $\bar \Y$. For a classification task, the cost is usually taken to be the cross-entropy
\begin{equation*}
	C(\Y, \bar \Y) = - \sum_{i=1}^{N_{pix}} \sum_{j=1}^{N_{classes}} \bar y_{ij} \log(y_{ij}),
\end{equation*}
where $\bar \Y \in \R^{N_{pix} \times N_{classes}}$ is the ground truth label indicator, i.e., $\bar y_{ij} = 1$ if pixel $i$ of sample $\X$ belongs to class $j$ and is zero otherwise. For global prediction, the cost is as if $N_{pix} = 1$.
For a regression task, a common choice is the mean squared error (MSE)
\begin{equation*}
	C(\Y, \bar \Y) = \frac{\|\Y-\bar{\Y}\|_2^2}{N_{pix} F_{out}} = \frac{1}{N_{pix} F_{out}} \sum_{i=1}^{N_{pix}} \sum_{j=1}^{F_{out}} (y_{ij} - \bar y_{ij})^2,
\end{equation*}
where $\bar{\Y}$ is the desired output. Again, take $N_{pix} = 1$ for global regression. We emphasize that the cost function and the SM layer is the sole difference between a neural network engineered for classification or regression.

The goal of learning is to find the parameters $\theta$ of the neural network that minimize the risk $R(\theta) = \esp \left[ C \left( NN_\theta(\X), \bar \Y \right) \right]$. In general, that expectation cannot be computed as the data distribution is unknown. We can however minimize an approximation, the empirical risk over the training set $\left\{ \left( \X_i, \bar \Y_i \right) \right\}_{i=1}^{N_{samples}}$:
\begin{equation*}
	\hat{\theta} = \argmin_\theta \sum_{i=1}^{N_{samples}} C \left(NN_\theta(\X_i), \bar \Y_i \right).
\end{equation*}
The optimization is performed by computing an error gradient w.r.t.\ all the parameters by back-propagation and updating them with a form of stochastic gradient descent (SGD):
\begin{equation*}
	\theta \leftarrow \theta - \frac{\eta}{|\B|} \sum_{i \in \B} \frac{\partial C \left( NN_\theta(\X_i), \bar \Y_i \right)}{\partial \theta} ,
\end{equation*}
where $\eta$ is the learning rate, and $\B$ is the set of indices in a mini-batch. Batches are used instead of single samples to gain speed by exploiting the parallelism afforded by modern computing platforms.
